DeconvOptim.jl - Signal Deconvolution with Julia

### 2021-07-09 paper/paper.tex
Update paper

Deconvolution is a versatile method to enhance the quality of signals measured with systems which can 
be expressed mathematically as a convolution of a system's response function with a signal. 
In this paper, we present \verb|DeconvOptim.jl|, a flexible toolbox written in Julia to deconvolve multi-dimensional signals which have been
degraded by a multi-dimensional signal response function. 
\verb|DeconvOptim.jl| works both on CPUs and GPUs and utilizes recent advancements in Julias automatic differentiation ecosystem.

@@

\section{Introduction}
Deconvolution has been a long addressed problem especially in astronomical imaging, digital signal processing or microscopical imaging.
The assumption is that images are degraded by some kind of blur which can be described as a convolution between a kernel $h$ and the image $S$:
\begin{equation}
    I(\mathbf r) = (h * S)(\mathbf r)
\end{equation}
$h$ is often called the point spread function (PSF) since it characterizes how the systems maps a point in the object to a blurred area
in the image.
Often this convolution operation is disturbed by different types of noise where the most common are additive Gaussian noise (e.g. due to the read out of the camera sensors) or Poisson shot noise (e.g. due to the quantum nature of photons):
\begin{align}
    Y_{\text{Poisson}}(\mathbf r) &= \text{Poisson}[(h * S)(\mathbf r)]\\
    Y_{\text{Gauss}}(\mathbf r) &= (h * S)(\mathbf r) + \text{Gauss}_{\mu, \sigma}(\mathbf r)
\end{align}
Due to noise and the low pass filter effect of the PSF, deconvolution is an ill-posed problem which cannot be solved directly.\\
A well known approach, because of its simplicity, has been the iterative Richardson-Lucy deconvolution algorithm \cite{Richardson:72}, \cite{lucy:74} for Poisson noise degraded signals.
Another common method is the non-iterative Wiener filter \cite{wiener2013extrapolation} because of its predictable runtime.
The Wiener filter assumes additive noise which is often not true, especially in low light conditions as in microscopy or astronomy. 
Recently machine learning based methods have become more and more favoured in deconvolution methods because
of the possibility to add learned properties of similar images to the algorithm \cite{Kruse_2017_ICCV}.
However, those implementations are usually restricted to cases where experimental datasets are available for the characterization and training.
\verb|DeconvOptim.jl| is a toolbox written in Julia \cite{bezanson2017julia} which offers several deconvolution algorithms like the iterative Richardson-Lucy algorithms but also
a much more flexible methods which are based on a loss function being minimized by a gradient descent.
Also this toolbox is not restricted in dimensionality and therefore can be applied to deconvolution problems in many different fields.

\section{Efficient Convolutions}
The mathematical definition of a convolution is
\begin{equation}
    (h * S)(\mathbf r) = \int_{-\infty}^{\infty} h(\mathbf r - \mathbf x) \cdot S(\mathbf x) \, \mathrm d \mathbf x.
    \label{eq:conv}
\end{equation}
To obtain efficient deconvolution algorithms we also need a efficient way to calculate the convolution operations.
In these times, convolution operations of spatially small kernels $h$ can be calculated on GPUs very efficiently.
However, often the blurring kernel $h$ is not spatially small and might be also multi-dimensional. In that case, FFT based convolutions
outperform sliding kernels. In microscopy and astronomy that is often the case, since a defocus broadens the PSF.

\subsection{FFT Based Convolution}
    The basis for fast convolution operations lies in the convolution theorem. It states that
    \begin{equation}
        \mathcal F\left\{(h * S)\right\}(\mathbf k) = \mathcal F\{h\} (\mathbf k) \cdot \mathcal F\{S\} (\mathbf k),
        \label{eq:fftconv}
    \end{equation} 
        which means that a Fourier transform of a  convolution operation is equivalent to a point wise multiplication of the Fourier transforms 
        of $h$ and $S$.
        Fortunately, the Fourier transform operations can be calculated even for moderate large 3D arrays quickly on modern computing machines via
        the Fast Fourier Transform (FFT).
        In Julia, \verb|FFTW.jl| offers a convenient interface to the FFTW library \cite{FFTW05}.
        The time complexity to calculate \autoref{eq:fftconv} is $\mathcal{O}(N \cdot \log N)$ for a one dimensional discretized dataset with $N$ samples.
        For a 2D dataset with size $N \times M$ it is $\mathcal{O}(N \cdot M \log(N + M))$ and accordingly for higher dimensions.

@@

\subsection{Wrap-Around Artifacts in FFT based Convolutions}
    One drawback of a FFT based convolution are wrap-around artifacts. The reason is, that the FFT operation calculates the discrete
    Fourier transform (DFT) which inherently assumes periodic data. If one applies a FFT on a finite data set the assumption is that
    this finite data set repeats itself. 
    To weaken the wrap-around artifacts in the convolution one can zero or mean pad the image before convolution 
    and remove the padding afterwards. Such approaches still suffer from artifacts when boundary values do not fit
    to the mean or to zero as it can be observed in \autoref{fig:withpad} at the top.
    \begin{figure}[h]
        \centering
        \begin{subfigure}[b]{.25\textwidth}
            \centering
            \begin{tikzpicture}
                \node at (0,0) {\includegraphics[width=.9\textwidth]{figures/full_house_no_pad.png}};
                \draw[red] (-2.0, -1.6) rectangle (-1.6, -1.2);
                \node[color=red] at (0,0) {\includegraphics[width=.4\textwidth]{figures/edge_house_no_pad.png}};
                \draw[color=red] (-0.91, -0.9) rectangle (0.9, 0.9); 
                \draw[color=red] (-2.0, -1.2) -- (-0.9, 0.9);
                \draw[color=red] (-1.6, -1.6) -- (0.9, -0.9);
            \end{tikzpicture}
            \caption{Without padding}
            \label{fig:withoutpad}
        \end{subfigure}%
        \begin{subfigure}[b]{.25\textwidth}
            \centering
            \begin{tikzpicture}
                \node at (0,0) {\includegraphics[width=.9\textwidth]{figures/full_house_pad.png}};
                \draw[red] (-2.0, -1.6) rectangle (-1.6, -1.2);
                \node[color=red] at (0,0) {\includegraphics[width=.4\textwidth]{figures/edge_house_pad.png}};
                \draw[color=red] (-0.90, -0.9) rectangle (0.9, 0.9); 
                \draw[color=red] (-2.0, -1.2) -- (-0.9, 0.9);
                \draw[color=red] (-1.6, -1.6) -- (0.9, -0.9);
            \end{tikzpicture}
            \caption{With padding.}
            \label{fig:withpad}
        \end{subfigure}
        \caption{Comparison between two blurred images. The image a) is a straightforward FFT based blurred image.
        The zoomed region shows dark artifacts at the left side originated from the right hand side. Also at the top of the image some dark regions from
        the bottom scattered into the bright sky.
        Image b) was padded before blurring with 10 pixels on each side with value 0.7. The blur kernel was a Gaussian kernel
        with standard deviation $\sigma=\SI{3}{pixel}$.}
        \label{fig:}
    \end{figure}
    To solve for the wrap-around artifacts in deconvolution we use a different approach which will be presented later. 
\subsection{Preferred Convolution}
The current state of \verb|DeconvOptim.jl| uses the FFT approach to calculate the convolution.
However, in principle one could add the possibility to change to spatially based convolutions.
But since the runtime gets especially long in multi-dimensional signal deconvolution where FFT based methods
are faster, we always use FFT based convolutions.

@@

At the beginning of the optimization we start with an initial guess $x(\mathbf r)$. 
Via a mapping function we can impose some restrictions to the final solution (e.g. non-negativity, certain value intervals).
The forward model is then applied to the mapped reconstruction. We provide a default function where 
the forward function is a convolution but we also allow to plug in a different forward model into the pipeline.
This output of the forward function is compared
with our noisy measurement under a certain loss function.
Since deconvolution is an ill-posed problem we regularize the loss function with a regularizer.
Afterwards we calculate the gradient of the loss function with respect to the reconstruction $x(\mathbf r)$ via automatic differentation (AD).
The values of the total loss and the gradient are then used by an optimizer (to minimize the total loss) to apply some kind of gradient descent optimization to get a proper reconstruction.\\

As typical for Julia packages, except \verb|FFTW.jl| and \verb|CUDA.jl| \cite{besard2018juliagpu}, all code is written in Julia.
Therefore it is easy to replace the forward model with any Julia code, as long as we can 
automatically differentiate through the code. The gradient step is currently calculated with the reverse mode AD of \verb|Zygote.jl| \cite{Zygote.jl-2018}. 
As default for the optimization we use the L-BFGS \cite{LBFGS} routine provided by \verb|Optim.jl| \cite{mogensen2018optim}.
L-BFGS is suitable for optimization with many million parameters as it only stores a sparse representation of the inverse Hessian matrix
which is built from a finite number of past function and gradient evaluations. 
In comparison to the Richardson-Lucy algorithm results are often good after 10 to 30 iterations whereas Richardson-Lucy takes
a few hundreds iterations. 
However, as we will see later due to the more complex optimization routine the speed-up in using L-BFGS is for a simple 
deconvolution minor.

\subsection{Mapping Function}
    Mapping functions are an easy way to constrain the resulting reconstruction to certain values.
    For example, if the optimizer optimizes the variables $\text{rec}$ which are then squared, we know that
    $\text{rec}^2$ is definitely non-negative. Instead of directly using $\text{rec}$ we compare
    $\text{rec}^2$ with the measurement $Y$. This mapping function does not change the outcome as long
    as it is an surjective function for the desired number space (e.g. a parabola is non-negative and covers all positive values) 
    and does not change the convexity of a problem.
    Functions like shifted $\tanh$ are possible as well since it is an bijective function in $[-1, 1]$ and monotonic increasing
    in the whole definition space.
    Hence the final result is $\text{rec}^2$ in case of the quadratic mapping.

\subsection{Wrap-around Artifacts Suppression}
    To prevent that wrap-around artifacts occur during deconvolution we can instead
    of reconstructing an object having the same size as the measurement, reconstruct
    an object having a slightly larger size. This is shown in \autoref{fig:wrap_around}.
    \begin{figure}[h]
        \centering
        \begin{tikzpicture}
            \draw[fill=red!20!white] (0,0) rectangle (4,4);
            \node[color=red] at (2, 3.8) {Padded area};
            \draw[fill=gray!40!white] (0.5,0.5) rectangle (3.5,3.5); 
            \node[text width=2.5cm, color=black] at (2, 2) {area compared with measurement};
        \end{tikzpicture}
        \caption{Reconstructed data is padded before deconvolution but only the red region is compared with the measurement.}
        \label{fig:wrap_around}
    \end{figure}
    Consequently, we only compare the grey part of the reconstruction with the measurement.
    Through the AD and the convolution operations we still optimize for the whole area (including the red one).
    With that padding we prevent that images on the boundary influence in circular manner the reconstruction.

\section{Loss Function}
    Besides the forward model, the loss function is the most critical part in the optimization.
    It has to take account for the noise created by the measurement process or device.

\subsection{Poisson Data Term}
    As already mentioned, Poisson shot noise is often the dominant source of noise hence we 
    explain the details.
    The key idea is to interpret the measurement as a stochastic process. Our aim is to find a deconvolved image which describes as accurate as possible the measured image.
    Mathematically the probability for a certain measurement $Y$ under Poisson noise is

@@

where $Y$ is the measurement, $\mu$ is the expected measurement (ideal measurement without noise) and $\Gamma$ is the generalized factorial function.
In the deconvolution process we get $Y$ as measured input and want to find the ideal specimen $S$ which results in a measurement $\mu(\mathbf r) = (S * \text{PSF})(\mathbf r))$.
Since we want to obtain for the best reconstruction, we want to find a $\mu(\mathbf r)$ so that $p(Y(\mathbf r) | \mu(\mathbf r))$ gets as large as possible. Because that means
that we find the specimen which describes the measurement with the highest probability.
Instead of maximizing $p(Y(\mathbf r) | \mu(\mathbf r))$ a common trick is to minimize $- \log(p(Y(\mathbf r)|\mu(\mathbf r)))$. 
Mathematically, the optimization of both functions provides same results but the latter is numerically more stable.

\begin{align}
&\underset{S(\mathbf r)}{\arg \min} (- \log(p(Y(\mathbf r)|\mu(\mathbf r)))) =\\ &\underset{S(\mathbf r)}{\arg \min} \sum_r (\mu(\mathbf r) + \log(\Gamma(Y(\mathbf r) + 1)) - Y(\mathbf r) \log(\mu(\mathbf r))
\end{align}

@@

since the second term only depends on $Y(\mathbf r)$ but not on $\mu(\mathbf r)$.
The gradient of $L$ with respect to $\mu(\mathbf r)$ is simply

@@

The function $L$ and the gradient $\nabla L$ are needed for any gradient descent optimization algorithm.
The numerical evaluation of the Poisson loss can lead to issues. Since $\mu(r)=0$ can happen for a measurement with zero intensity background. However, the loss is not defined for $\mu \leq 0$. In our source code we set all intensity values below a certain threshold $\epsilon$ to $\epsilon$ itself. This prevents the evaluation of the logarithm at undefined values.
The final source code has a modified version of \autoref{eq:gradpoisson} as registered gradient for \autoref{eq:poisson_loss}.
The reason is, that for values of $\mu(\mathbf r) = 0$ the equation is not defined and has to be handled separately.

@@

    \subsection{Influence of Regularizer}
        In this part we want to demonstrate the influence of the regularizer.
        \autoref{fig:fabioa} is a simulated imaged which has been affected by Poisson shot noise and blurring by a PSF.
        \begin{figure}[h]
            \begin{subfigure}[b]{.25\textwidth}
                \centering
                \includegraphics[width=\textwidth]{figures/resolution_test_chart_blurry.png}
                \caption{Measured with Poisson noise}
                \label{fig:fabioa}
            \end{subfigure}%
            \hfill
            \begin{subfigure}[b]{.25\textwidth}
                \centering
                \includegraphics[width=\textwidth]{figures/resolution_test_chart_no_reg_iter_50.png}
                \caption{Without regularizer $N=50$}
                \label{fig:fabiob}
            \end{subfigure}%
            \hfill
            \begin{subfigure}[b]{.25\textwidth}
                \centering
                \includegraphics[width=\textwidth]{figures/resolution_test_chart_no_reg_iter_10.png}
                \caption{Without regularizer $N=10$}
                \label{fig:fabioc}
            \end{subfigure}%            
            \hfill
            \begin{subfigure}[b]{.25\textwidth}
                \centering
                \includegraphics[width=\textwidth]{figures/resolution_test_chart_regularizer_008_iter_21.png}
                \caption{With regularizer $N=21$.}
                \label{fig:fabiod}
            \end{subfigure}%            
            \caption{a) has been convolved with a PSF and affected by Poisson shot noise (100 photons expected for the brightest pixel). 
                        Deconvolved b) for 50 iterations without a regularizer c) for 10 iterations without a regularizer d) with Good's roughness regularizer and 
                            converged after 21 iterations with $\lambda=0.08$.}
            \label{fig:image}
        \end{figure}

@@

        are getting closer in the center which is equivalent to higher spatial frequencies.

@@

        Additionally, we notice that the core in the Siemens star is smaller meaning better spatial resolution. 

@@

        \autoref{fig:fabiod} shows the results after 21 iterations of the deconvolution but regularized with a Good's roughness regularizer.

@@

        resolution. With the regularizer weight $\lambda$ one can adjust the strength.

@@

        this frequency could not be recovered at all, $G_\textrm{R} < 0$ means that the reconstruction introduces fake frequencies
        which do not fit to the ground truth at all.

@@

                In a) we show the REG for the hand-optimized non-regularized version with 10 iterations and the (converged) regularized versions with $\lambda=0.08$ and $\lambda=0.5$.}

@@

        The green curve shows the REG of the deconvolution without a regularizer. For certain regions below  $f_\textrm{max}$

@@

        The regularized blue curve performs better and only certain frequencies show a REG below zero.

@@

        Especially the deconvolution without an regularizer suffers from poor performance after a high number of iterations.

@@

        The regularized version seems to be robust against a higher number of iterations and even converges after 21.

@@

    \subsection{Multi-Dimensional Microscopy Data}
        To show the full ability, we deconvolve a four dimensional dataset where the measured volume consists of a 3D datasets
        with three different color channels which is shown in \autoref{fig:result_3d}.
        The PSF is only three dimensional.
        Julia's broadcast mechanism handles that the PSF is only three dimensional and therefore
        we only need to specify over which dimensions the convolution should happen.
        \begin{figure}[h]
            \begin{subfigure}{.25\textwidth}
                \centering
                \includegraphics[width = .95\textwidth]{figures/raw_rgb.png}
            \end{subfigure}%
            \begin{subfigure}{.25\textwidth}
                \centering
                \includegraphics[width = .95\textwidth]{figures/deconvolved_rgb.png}
            \end{subfigure}
            \caption{Slice of a 3D dataset with three different colors. a) is the raw data and b) is the
                    deconvolved image.}
            \label{fig:result_3d}
        \end{figure} 
        Clearly we can see that the deconvolution removed a lot of the background blur.

>>>>

Deconvolution is a versatile method to enhance the quality of signals measured with systems which can 
be expressed mathematically as a convolution of a system's response function with a signal. 
In this paper, we present \verb|DeconvOptim.jl|, a flexible toolbox written in Julia to deconvolve one or multiple multi-dimensional signals which have been
degraded by a multi-dimensional signal response function. 
\verb|DeconvOptim.jl| works both on CPUs and GPUs and utilizes recent advancements in Julias automatic differentiation ecosystem.

@@

\todo{multi view deconvolution}

\section{Introduction}
Deconvolution has been a long addressed problem especially in astronomical imaging, digital signal processing or microscopy imaging.
The assumption is that images are degraded by some kind of blur which can be described as a convolution of a kernel $h$ with the image $S$:
\begin{equation}
    I(\mathbf r) = (h * S)(\mathbf r)
\end{equation}
where $*$ denotes a convolution operation.
$h$ is often called the point spread function (PSF) since it characterizes how the systems maps a point in the object to a blurred spot 
in the image.
Often this convolution operation is disturbed by different types of noise where the most common are additive Gaussian noise (e.g. due to the read out of the camera sensors) or Poisson shot noise (e.g. due to the quantum nature of photons):
\begin{align}
    Y_{\text{Poisson}}(\mathbf r) &= \text{Poisson}[(h * S)(\mathbf r)]\\
    Y_{\text{Gauss}}(\mathbf r) &= (h * S)(\mathbf r) + \text{Gauss}_{\sigma}(\mathbf r)
\end{align}
Due to noise and the low pass filter effect of the PSF, deconvolution is an ill-posed problem which cannot be solved directly.\\
A well known approach, because of its simplicity, has been the iterative Richardson-Lucy deconvolution algorithm \cite{Richardson:72}, \cite{lucy:74} for Poisson noise degraded signals.
Another common method is the non-iterative Wiener filter \cite{wiener2013extrapolation} because of its predictable runtime.
The Wiener filter assumes additive noise which is often not a good approximation, especially in low light conditions as in microscopy or astronomy. 
Recently machine learning based methods have become more and more favoured in deconvolution because
of the possibility to add learned properties of similar images to the algorithm \cite{Kruse_2017_ICCV}.
However, those implementations are usually restricted to cases where experimental datasets are available for the characterization and training.
\verb|DeconvOptim.jl| is a toolbox written in Julia \cite{bezanson2017julia} which offers several deconvolution algorithms like the iterative Richardson-Lucy algorithms but also
much more flexible methods which are based on minimizing a 
noise-dependent loss function with the help of modern gradient-descent minimization routines. 
Also this toolbox is not restricted in dimensionality and therefore can be applied to deconvolution problems in many different fields.

\section{Efficient Convolutions}
The mathematical definition of a convolution is
\begin{equation}
    (h * S)(\mathbf r) = \int_{-\infty}^{\infty} h(\mathbf r - \mathbf x) \cdot S(\mathbf x) \, \mathrm d \mathbf x.
    \label{eq:conv}
\end{equation}
To obtain efficient deconvolution algorithms we also need a efficient way to calculate the convolution operations.
Today, convolution operations of spatially small kernels $h$ can be calculated on GPUs very efficiently in real space.
However, often the blurring kernel $h$ is not spatially small and might be also multi-dimensional. In that case, FFT based convolutions
outperform sliding kernels. In microscopy and astronomy that is often the case, since a defocus broadens the PSF.

\subsection{FFT Based Convolution}
    The basis for fast convolution calculation via Fourier transforms lies in the convolution theorem. It states that
    \begin{equation}
        \mathcal F\left\{(h * S)\right\}(\mathbf k) = \mathcal F\{h\} (\mathbf k) \cdot \mathcal F\{S\} (\mathbf k),
        \label{eq:fftconv}
    \end{equation} 
        which means that a Fourier transform of a  convolution operation is equivalent to a point wise multiplication of the Fourier transforms 
        of $h$ and $S$.
        Fortunately, the Fourier transform operations can be calculated even for moderate large 3D arrays quickly on modern computing machines via
        the Fast Fourier Transform (FFT).
        In Julia, \verb|FFTW.jl| offers a convenient interface to the FFTW library \cite{FFTW05}.
        The time complexity to calculate \autoref{eq:fftconv} is $\mathcal{O}(N \cdot \log N)$ for a one dimensional discretized dataset with $N$ samples.
        For a 2D dataset with size $N \times M$ it is $\mathcal{O}(N \cdot M \log(N \cdot M))$ and accordingly for higher dimensions.

@@

\subsection{Wrap-Around Artifacts in FFT based Convolutions}
    One drawback of a FFT based convolution are wrap-around artifacts (see arrow in \autoref{fig:withoutpad}). The reason is, that the FFT operation calculates the discrete
    Fourier transform (DFT) which inherently assumes periodic data. If one applies a FFT on a finite data set the assumption is that
    this finite data set repeats itself. 
    To weaken the wrap-around artifacts in the convolution one can zero or mean pad the image before convolution 
    and remove the padding afterwards. Such approaches still suffer from artifacts when boundary values do not fit
    to the mean or to zero as it can be observed in \autoref{fig:withpad} where the arrow is pointing to.
    \begin{figure}[h]
        \centering
        \begin{subfigure}[b]{.25\textwidth}
            \centering
            \begin{tikzpicture}
                \node at (0,0) {\includegraphics[width=.9\textwidth]{figures/full_house_no_pad.png}};
                \draw[red] (-2.0, -1.6) rectangle (-1.6, -1.2);
                \node[color=red] at (0,0) {\includegraphics[width=.4\textwidth]{figures/edge_house_no_pad.png}};
                \draw[color=red] (-0.91, -0.9) rectangle (0.9, 0.9); 
                \draw[color=red] (-2.0, -1.2) -- (-0.9, 0.9);
                \draw[color=red] (-1.6, -1.6) -- (0.9, -0.9);
                \draw[<-, color=red, thick ](-0.85, 0) -- (-0.4, 0.6);
            \end{tikzpicture}
            \caption{Without padding}
            \label{fig:withoutpad}
        \end{subfigure}%
        \begin{subfigure}[b]{.25\textwidth}
            \centering
            \begin{tikzpicture}
                \node at (0,0) {\includegraphics[width=.9\textwidth]{figures/full_house_pad.png}};
                \draw[red] (-2.0, -1.6) rectangle (-1.6, -1.2);
                \node[color=red] at (0,0) {\includegraphics[width=.4\textwidth]{figures/edge_house_pad.png}};
                \draw[color=red] (-0.90, -0.9) rectangle (0.9, 0.9); 
                \draw[color=red] (-2.0, -1.2) -- (-0.9, 0.9);
                \draw[color=red] (-1.6, -1.6) -- (0.9, -0.9);
                \draw[<-, color=red, thick ](-0.8, -0.7) -- (-0.3, -0.3);
            \end{tikzpicture}
            \caption{With padding.}
            \label{fig:withpad}
        \end{subfigure}
        \caption{Comparison between two blurred images. The image a) is a straightforward FFT based blurred image.
        The zoomed region shows dark artifacts at the left side originated from the right hand side. Also at the top of the image some dark regions from
        the bottom scattered into the bright sky.
        Image b) was padded before blurring with 10 pixels on each side with value 0.7 which results in bright regions around the image. The blur kernel was a Gaussian kernel
        with standard deviation $\sigma=\SI{3}{pixel}$.}
        \label{fig:}
    \end{figure}
    To solve for the wrap-around artifacts in deconvolution we use a different approach which will be presented in \autoref{sec:wrap}. 
\subsection{Preferred Convolution}
The current state of \verb|DeconvOptim.jl| uses the FFT approach to calculate the convolution.
However, in principle one could add the possibility to choose spatially based convolutions.
But since the runtime gets especially long in multi-dimensional signal deconvolution where FFT based methods
are faster, we always use FFT based convolutions here.

@@

At the beginning of the optimization we start with an initial guess $x(\mathbf r)$. 
Via a mapping function we can impose some restrictions to the final solution (e.g. non-negativity, value intervals).
The forward model is then applied to the mapped reconstruction. We provide a default function where 
the forward function is a convolution but we also allow to plug in a different forward model.
This output of the forward function is compared
to our noisy measurement under a given loss function.
Since deconvolution is an ill-posed problem, we regularize the loss function with a regularizer.
Afterwards we calculate the gradient of the loss function with respect to the reconstruction $x(\mathbf r)$ via automatic differentation (AD).
The values of the total loss and the gradient are then used by an optimizer (to minimize the total loss) to apply a gradient descent optimization to get a proper reconstruction.\\

As typical for Julia packages, except \verb|FFTW.jl| and \verb|CUDA.jl| \cite{besard2018juliagpu}, all code is written in Julia.
Therefore it is easy to replace the forward model with any Julia code, as long as we can 
automatically differentiate through the code. The gradient step is currently calculated with the reverse mode AD of \verb|Zygote.jl| \cite{Zygote.jl-2018}. 
As default for the optimization we use the L-BFGS \cite{LBFGS} routine provided by \verb|Optim.jl| \cite{mogensen2018optim}.
L-BFGS is suitable for optimization with many million parameters as it only stores a sparse representation of the inverse Hessian matrix,
which is built from a finite number of past function and gradient evaluations. 
Results are often good after 10 to 30 iterations whereas Richardson-Lucy takes
a few hundreds iterations. 
However, as we will see later due to the more complex optimization routine the speed-up in using L-BFGS is for a simple 
deconvolution minor.

\subsection{Mapping Function}
    Mapping functions are an easy way to constrain the resulting reconstruction to predefined ranges.
    For example, if the optimizer optimizes the variables $\text{rec}$ which are then squared, we know that
    $\text{rec}^2$ is definitely non-negative. Instead of directly using $\text{fwd}(\text{rec})$ we compare
    $\text{fwd}(\text{rec}^2)$ with the measurement $Y$. This mapping function does not change the outcome as long
    as it is an surjective function for the desired number space (e.g. a parabola is non-negative and covers all positive values) 
    and does not change the convexity of a problem.
    Functions like shifted $\tanh$ are possible as well since it is an bijective function in $[-1, 1]$ and monotonic increasing
    in the whole definition space.
    Hence the final result is $\text{rec}^2$ in case of the quadratic mapping.

\subsection{Wrap-around Artifacts Suppression}
    \label{sec:wrap}
    To prevent that wrap-around artifacts occur during deconvolution we can instead
    of reconstructing an object having the same size as the measurement, reconstruct
    an object having a slightly larger size. This is shown in \autoref{fig:wrap_around}.
    \begin{figure}[h]
        \centering
        \begin{tikzpicture}
            \draw[fill=red!20!white] (0,0) rectangle (4,4);
            \node[color=red] at (2, 3.8) {Padded area};
            \draw[fill=gray!40!white] (0.5,0.5) rectangle (3.5,3.5); 
            \node[text width=2.5cm, color=black] at (2, 2) {area compared with measurement};
        \end{tikzpicture}
        \caption{Reconstructed data is padded before deconvolution but only the red region is compared with the measurement.}
        \label{fig:wrap_around}
    \end{figure}
    Consequently, we only compare the grey part of the reconstruction with the measurement.
    Through the AD and the convolution operations we still optimize for the whole area (including the red one).
    With that padding we reduce that images on the boundary influence in circular manner on reconstruction in the grey area.

\section{Loss Function}
    Apart from forward model, the loss function is the most critical part in the optimization.
    It has to account for the noise created by the measurement process or device.

\subsection{Poisson Data Term}
    Poisson shot noise is often the dominant source of noise hence we 
    explain the details.
    The key is to interpret the measurement as a stochastic process. 
    Our aim is to find a deconvolved image which predicts the measured image as accurately as 
    possible.
    Mathematically the probability for a certain measurement $Y$ under Poisson noise is

@@

where $Y$ is the measurement, $\mu$ is the expected measurement (ideal measurement without noise) and $\Gamma$ is the generalized factorial function.
In the deconvolution process we get $Y$ as measured input and want to find the ideal specimen $S$ which results in a measurement $\mu(\mathbf r) = (S * \text{PSF})(\mathbf r)$.
Since we want to obtain the best reconstruction, we want 
to find a $\mu(\mathbf r)$ so that $p(Y(\mathbf r) | \mu(\mathbf r))$ gets maximized.
Because that means
that we find the specimen which describes the measurement with the highest probability.
Instead of maximizing $p(Y(\mathbf r) | \mu(\mathbf r))$ a common trick is to minimize $- \log(p(Y(\mathbf r)|\mu(\mathbf r)))$. 
Mathematically, the optimization of both functions provides same results but the latter is numerically more stable.

\begin{align}
&\underset{S(\mathbf r)}{\arg \min} (- \log(p(Y(\mathbf r)|\mu(\mathbf r)))) =\\ &\underset{S(\mathbf r)}{\arg \min} \sum_r \mu(\mathbf r) + \log(\Gamma(Y(\mathbf r) + 1)) - Y(\mathbf r) \log(\mu(\mathbf r))
\end{align}

@@

since the second term only depends on the constant $Y(\mathbf r)$ but not on $\mu(\mathbf r)$.
The gradient of $L$ with respect to $\mu(\mathbf r)$ is simply

@@

The function $L$ and the gradient $\nabla L$ are needed for any gradient descent optimization algorithm.
The numerical evaluation of the Poisson loss can lead to issues. Since $\mu(r)=0$ can happen for a measurement with zero intensity background. However, the loss is not defined for $\mu \leq 0$. In our source code we set all intensity values below a predefined threshold $\epsilon$ to $\epsilon$ itself. This prevents the evaluation of the logarithm at undefined values.
The final source code has a modified version of \autoref{eq:gradpoisson} as registered gradient for \autoref{eq:poisson_loss}.
The reason is, that for values of $\mu(\mathbf r) = 0$ the equation is not defined and has to be handled separately.

@@

    \subsection{Influence of Regularizer}
        In this part we want to demonstrate the influence of the regularizer.
        \autoref{fig:fabioa} is a simulated imaged which has been affected by Poisson shot noise and blurring by a PSF.
        \begin{figure}[h]
            \begin{subfigure}[b]{.25\textwidth}
                \centering
                \includegraphics[width=\textwidth]{figures/resolution_test_chart_blurry.png}
                \caption{Measured with Poisson noise}
                \label{fig:fabioa}
            \end{subfigure}%
            \hfill
            \begin{subfigure}[b]{.25\textwidth}
                \centering
                \includegraphics[width=\textwidth]{figures/resolution_test_chart_no_reg_iter_50.png}
                \caption{Without regularizer $N=50$}
                \label{fig:fabiob}
            \end{subfigure}%
            \hfill
            \begin{subfigure}[b]{.25\textwidth}
                \centering
                \includegraphics[width=\textwidth]{figures/resolution_test_chart_no_reg_iter_10.png}
                \caption{Without regularizer $N=10$}
                \label{fig:fabioc}
            \end{subfigure}%            
            \hfill
            \begin{subfigure}[b]{.25\textwidth}
                \centering
                \includegraphics[width=\textwidth]{figures/resolution_test_chart_regularizer_008_iter_21.png}
                \caption{With regularizer $N=24$.}
                \label{fig:fabiod}
            \end{subfigure}%            
            \caption{a) has been convolved with a PSF and affected by Poisson shot noise (100 photons expected for the brightest pixel). 
                        Deconvolved b) for 50 iterations without a regularizer c) for 10 iterations without a regularizer d) with Good's roughness regularizer and 
                            converged after 24 iterations with $\lambda=0.03$.}
            \label{fig:image}
        \end{figure}

@@

        have decreasing spacing towards its middle which is equivalent to higher spatial frequencies.

@@

        Additionally, we notice that the core in the Siemens star is denser is denser 
        requiring higher spatial resolution to resolve. 

@@

        \autoref{fig:fabiod} shows the results after 24 iterations of the deconvolution but regularized with a Good's roughness regularizer.

@@

        resolution. With the regularizer weight $\lambda$ one can adjust the regularization strength.

@@

        this frequency could not be recovered at all, $G_\textrm{R} < 0$ means that the reconstruction introduces 
        wrong information at this frequencies
        which are further from the ground truth then not reconstructing these frequencies. 

@@

                In a) we show the REG for the hand-optimized non-regularized version with 10 iterations and the (converged) regularized versions with $\lambda=0.03$ and $\lambda=0.5$.}

@@

        The green curve shows the REG of the deconvolution without a regularizer. 
        For regions below  $f_\textrm{max}$

@@

        The regularized blue curve performs better and only some frequencies show a REG below zero.

@@

        Especially the deconvolution without a regularizer suffers from poor performance after a high number of iterations.

@@

        The regularized version seems to be robust against a higher number of iterations and even converges after 24.

@@

    \subsection{Multi-Dimensional Microscopy Data}
        To more completely demonstrate the features available in \verb|DeconvOptim.jl|,
        we deconvolve a four dimensional dataset where the measured volume consists of a 3D datasets
        with three different color channels, shown in \autoref{fig:result_3d}.
        Here we assumed a common single 3D PSF for all color channels.
        Julia's broadcast mechanism handles that the PSF is only three dimensional and therefore
        we only need to specify over which dimensions the convolution should happen.
        \begin{figure}[h]
            \begin{subfigure}{.25\textwidth}
                \centering
                \includegraphics[width = .95\textwidth]{figures/raw_rgb.png}
            \end{subfigure}%
            \begin{subfigure}{.25\textwidth}
                \centering
                \includegraphics[width = .95\textwidth]{figures/deconvolved_rgb.png}
            \end{subfigure}
            \caption{Slice of a 3D dataset with three different colors. a) is the raw data and b) is the
                    deconvolved image.}
            \label{fig:result_3d}
        \end{figure} 
        Clearly we can see that the deconvolution removed a lot of the background blur.
